---
title: "Live Session Assignment 2"
author: "Stuart Miller"
date: "January 8, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gridExtra)
library(tswge)

data1 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTest.csv')
data2 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTrain.csv')
data <- rbind(data2, data1)

data['seq'] <- seq_along(data[, 'humidity'])
```

## Assignment 2

1. Generate the spectral density for your realization.

```{r, spec}
series <- data %>% ggplot(aes(y = humidity, x = seq)) +
  geom_line() +
  xlab('Time (Sequence Number)') +
  ylab('Humidity')

spectrum <- as.data.frame(parzen.wge(data[, 'humidity'], plot = F))

spec.dens <- spectrum %>% ggplot(aes(y = pzgram, x = freq)) +
  geom_line() +
  xlab('Frequency') +
  ylab('Component Magnitude')

grid.arrange(series, spec.dens)

```

2. Indicate which frequencies, if any, appear to be in your series and/or if there is evidence of wandering behavior.

There are no clear frequencies present based on the spectral density plot. The sample frequency spectrum starts high at 0, then quickly decrease in magnitude in the first few frequency components, followed by a steady decrease as the frequency increases. This is indicative of aperiodic or wandering behavior.

3. Comment on how the information from your spectral density is relevant to your series. (For example, if the data is seasonal on an annual basis and recorded on a monthly basis, then you would expect to see a peak in the spectral density at 1/12 = .083.)

The resulting spectral density plot is not surprising. Visually, the realization appears to have some period nature, which is expected since it is weather data (typically seasonality is observed). However, this period, approximately 1/375 = 0.0027, is very close to zero.  It could be the case that more samples would be necessary to provide sufficient resolution for the frequency component to diverge from zero. 