---
title: "Live Session Assignments"
author: "Stuart Miller"
date: "January 2, 2020"
output: github_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tswge)
```

## Assignment 1

1. For live session, find an example of time-series data. Write a brief one- to three-sentence description.

I chose a climate dataset from kaggle.
The data can be downloaded from Kaggle [here](https://www.kaggle.com/sumanthvrao/daily-climate-time-series-data).

This is a time series of humidity from Delhi, India taken from 01/01/2013 to 04/24/2017.

```{r}
data1 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTest.csv')
data2 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTrain.csv')
data <- rbind(data2, data1)

data['seq'] <- seq_along(data[, 'humidity'])

data %>% ggplot(aes(y = humidity, x = seq)) +
  geom_line() +
  xlab('Time (Sequence Number)') +
  ylab('Humidity')
```

2. In your data set, identify  x1, x2, and x3.

The first three data points are

84.5, 92.0, and 87.0

```{r}
data[seq(1,3), 'humidity']
```


3. Using your time series that you selected, would you be able to obtain another realization?

This is climate data from a specific city over a specific time period.
So it would not be possible to obtain another realization.

4. With respect to your time series, does it appear that the assumption of constant mean of the distribution of each Xt is reasonable? 

The mean does appear to be constant.
There is (possibly seasonal) movement, but it appears to oscillate over a constant mean.

5. With respect to your time series, does it appear that the assumption of constant variance of the distribution of each Xt is reasonable?

The assumption of constant mean seems reasonable.
The variation along the movement does not seem to vary with time.

6. Provide the ACF for your time series. Is your time series stationary?

The ACF plot for the whole time series is shown in the top image. 
The ACF falls off quickly in the first few lags, then falls off gradually as the lags increase.
The data was split into 4 equal length segments and the ACF plots were produced (bottom). These 4 plots exhibit fairly similar characteristics. 
There is only a small amount of deviation between the 4 ACF plots.

There does not appear to be strong evidence that this time series is not stationary.


```{r}
acf(data[, 'humidity'])

ts <- data[, 'humidity']
ts.split <- split(ts, ceiling(seq_along(ts)/(length(ts)/4)))

par(mfrow=c(2,2))
acf(ts.split[[1]])
acf(ts.split[[2]])
acf(ts.split[[3]])
acf(ts.split[[4]])
```










