---
title: "Unit 3"
author: "Stuart Miller"
date: "January 15, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(tidyverse)
```


```{r}
Realization = gen.sigplusnoise.wge(200,coef = c(5,0),freq = c(.1,0), vara = 10, sn = 1)
ma = stats::filter(Realization,rep(1,5))/5
plot(ma)
```

1. Load the Data

```{r}

Walmart = read.csv('./Walmart.csv',header = TRUE)
Stor8Item1 = Walmart %>% dplyr::filter(item == 1 & store == 8)
```

2. Does the data come from a stationary process? 
Make sure and address all 3 conditions.

The realization is plotted below.
There appears to be a periodic oscialliation seasonaly.
This would indicate that the mean is time dependent, violating the assumption for stationarity.
Additionally, there appears to be a slight trend.


The variance appears to be constant over the movement.
There does not appear to be sufficient evidence to suggest that the assumption of constant variance is not met.


To assess the assuption of constant autocovariance, the data was plot into 4 segments and the ACF for these segments were plotted.
The ACF are similar, thuse there does not appear to be sufficient evidence to suggest that the assumption of autocorrelation is not met.

Since the assumption of constant mean appears to be violated,
we conclude that there is sufficient evidence to suggest that this process is **not stationary**.

```{r}
plot(Stor8Item1$sales)
acf(Stor8Item1$sales)
```

```{r}
ts <- Stor8Item1$sales
ts.split <- split(ts, ceiling(seq_along(ts)/(length(ts)/4)))

par(mfrow=c(2,2))
acf(ts.split[[1]])
acf(ts.split[[2]])
acf(ts.split[[3]])
acf(ts.split[[4]])
```


3. The client believes there is yearly and weekly seasonality in the data.  
Is there evidence of this with respect to the spectral density? 

For a weekly trend, we would expect to see evidence of a frequency at 1/7 = 0.143.
From the spectral density plot, evidence of this freqency exists.

For a weekly trend, we would expect to see evidence of a frequency at 1/365 = 0.0027.
This is very close to 0, we see a peak close to zero, but the year frequency is not distinushable from 0.

```{r}
parzen.wge(ts)
```

Applying a 5-point moving avarge filter to the sales data significantly smoothes out the variation in the data.
A moving average filter is a low pass filter.
A plot of the realization after filter shows a suppression in the high frequencies.
There is still some evidence of a spectral peak at 0.14 (the weekly trend), but is suppressed.

```{r}
ma.5p <- stats::filter(ts, rep(1,5)/5)
plot(ma.5p)
parzen.wge(ma.5p[seq(from = 3, to = length(ma.5p)-2)])
```

The 51 point moving average filter is a much stronger low pass filter.
The high frequeny is very suppressed.
Plotting the spectral density, All the high frequency appears to be filtered out.

```{r}
ma.51p <- stats::filter(ts, rep(1,51)/51)
plot(ma.51p)
parzen.wge(ma.51p[seq(from = 26, to = length(ma.51p)-26)])
```

