---
title: "Live_Session_Assignment_9"
author: "Stuart Miller"
date: "February 29, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
data <- read.csv('../texasgasprice.csv')
```

# Live Session Assignment 9


## Fit an AR(2) to the data using the maximum likelihood estimates like you did in the Concept Check question.

```{r}
# fit an AR2 with max likelihood est
est.mle <- est.ar.wge(data$Price)
```


## Fit and AR(2) to the data using the Burg estimates.  Display and describe.

```{r}
# fit an AR2 with max likelihood est
est.burg <- est.ar.wge(data$Price, type = 'burg')
```

## Find the ASE for the maximum likelihood fit by forecasting the last 24 weeks of the series.

```{r}
fore.mle <- fore.arma.wge(data$Price, phi = est.mle$phi, 
                          n.ahead = 24, lastn = T)
fore.mle.mse <- mean( (data$Price[(205-24+1):205] - fore.mle$f)^2 )
fore.mle.mse
```


## Find the ASE for the Burg fit by forecasting the last 24 weeks of the series.

```{r}
fore.burg <- fore.arma.wge(data$Price, phi = est.burg$phi,
                           n.ahead = 24, lastn = T)
fore.burg.mse <- mean( (data$Price[(205-24+1):205] - fore.burg$f)^2 )
fore.burg.mse
```

## Which model would you choose?


The two model appear to have about the same performance on the data. 
I do not see any reason to chose one over the other. 
I will chose the ML estimate in this case as the default option.