---
title: "Live_Session_Assignment_5"
author: "Stuart Miller"
date: "February 1, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tswge)
```

# Select a Model for the Weather Data

A stationary model will be selected for the weather data based on aic (`aic5.wge`).

**Replot the data**

```{r}
data1 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTest.csv')
data2 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTrain.csv')
data <- rbind(data2, data1)

data['seq'] <- seq_along(data[, 'humidity'])

data %>% ggplot(aes(y = humidity, x = seq)) +
  geom_line() +
  xlab('Time (Sequence Number)') +
  ylab('Humidity')
```

The best model based on AIC is an ARMA(4,1).

```{r}
aic5.wge(data$humidity)
```


# Select an ARMA model for the Walmart Data

Replot the data.

```{r}
Walmart = read.csv('./Walmart.csv',header = TRUE)
Stor9Item50 = Walmart %>% dplyr::filter(item == 50 & store == 9)
plotts.sample.wge(Stor9Item50$sales)
```



```{r}
aic5.wge(Stor9Item50$sales)
```


# Find $\rho1$

Find $\rho_1$ by hand for the following model

$$
X_t = a_t - 0.8 a_{t-1} + 0.5 a_{t-2}
$$

**Solution**

$$
\rho_1 = \frac{-\theta_1 + \theta_1 \theta_2}{1 + \theta_1 + \theta_2}
$$

For the model, $\theta_1 = 0.8$ and $\theta_2 = -0.5$.
Plugging in the values,

$$
\rho_1 = \frac{-0.8 + (0.8)(-0.5)}{1 + 0.8 - 0.5} = -0.923
$$

**R calculation for autocorrelation**

```{r}
(-.8 + 0.8*(-.5)) / (1 + 0.8 -0.5)
```

# Show the GLP form of a MA model

Convert the model below into GLP form

$$
X_t = a_t - 0.8 a_{t-1} + 0.5 a_{t-2}
$$

$$
X_t = \sum_{k=0}^{2} \psi_k a_{t-k} | \psi = \{ 1, -0.8, 0.5 \}
$$


```{r}
psi.weights.wge(theta = c(0.8, -0.5), lag.max = 2 )
```

# Generate ARMA Relatizations

The model from before.

```{r}
gen1 = gen.arma.wge(n = 100, phi = 0, theta = c(0.8, -0.5), plot = F)
gen2 = gen.arma.wge(n = 100, phi = c(0.8, -0.5), plot = F)
gen3 = gen.arma.wge(n = 100, phi = c(0.5,-0.6,0.5), theta = c(0.8, -0.5), plot = F)
```



```{r}
plotts.sample.wge(gen1)
```


```{r}
plotts.sample.wge(gen2)
```




```{r}
plotts.sample.wge(gen3)
```

# Find a model for the SWA data

Use AIC 5 to identify the top five quality models with respect to AIC for the cancelled flight data from the SWADelay.csv data set (column: arr_cancelled).
Comment on which are AR, MA, and ARMA.

```{r}
SWA <- read_csv('../swadelay.csv')
plotts.wge(SWA$arr_cancelled)
```

```{r}
plotts.sample.wge(SWA$arr_cancelled)
```


Find the best fit model with `aic5.wge`.

```{r}
aic5.wge(SWA$arr_cancelled)
```














































