---
title: "Untitled"
author: "Stuart Miller"
date: "February 15, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tswge)

# read in data
data1 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTest.csv')
data2 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTrain.csv')
data <- rbind(data2, data1)

data['seq'] <- seq_along(data[, 'humidity'])
```

## Assignment 1

Which model do you think is appropriate to use to forecast your time series? Why?

```{r}
plotts.sample.wge(data$humidity)
```

Looking at the parzen window with additional truncation points to see where most of the low frequency is positioned.

According to the parzen window, the maximum value of the spectral density (19) occurs at 0.005, which maps to a period of 200 days.
However, this period appears to be too small based on visual inspection of the realization.
Approximately, 375 days appears to be a better estimate of the period.
The spectral magnitude of the frequeny with period of 375 days is approximately 17, which is close to the maximum value of the spectral density.

```{r}
parzen.values <- plotts.parzen.wge(data$humidity, m2 = c(500, 1000))
print('max value of the parzen window occurs at freq')
parzen.values$freq[which.max(parzen.values$dbz2)]
print('the spectral magnitude at a period of 375 is')
parzen.values$dbz2[5]
```

I will use the seasonal model suggested by the simple frequency analysis and the model I chose based on visual inspection.
Based on validation, the model I would chose the model with s = 375.

```{r}
mse <- function(y.true, y.pred){
  mean( (y.true - y.pred)^2 )
}

fcast.375 <- fore.aruma.wge(data$humidity, s = 375, n.ahead = 750, lastn = T)
fcast.200 <- fore.aruma.wge(data$humidity, s = 200, n.ahead = 750, lastn = T)

test.set <- data$humidity[(length(data$humidity) - 750 + 1):length(data$humidity)]

mse(test.set, fcast.375$f)
mse(test.set, fcast.200$f)
```

The image below shows the forward forcase for 1000 time steps.

```{r}
fcast <- fore.aruma.wge(data$humidity, s = 375, n.ahead = 1000)
plot(fcast$f)
```

I removed the seasonality of s= 375 and checked to see if the resulting noise could be fit to an ARMA model.
Based on the r esults of `aic5`, the noise could be fit to an ARMA model.

```{r}
diffed.data <- artrans.wge(data$humidity, phi.tr = c(rep(0,374),1))
aic5.wge(diffed.data)
```



## Assignment 2

Find the first 5 psi weights for the model below, and use them to find the half-width of the 95 percentile probability interval for the third forecast $(\hat{X}(3))$.  Please show your work as well as a plot of the series and the first eight forecasts with probability intervals.

$$
(1–0.9B)(1–0.8B)X_t = a_t
$$

**Realization**

$X_1$ = 5, $X_2$ = 8, $X_3$ = 9, $X_4$ = 8,
$X_5$ = 7, $X_6$ = 6, $X_7$ = 4, $X_8$ = 3


```{r}
mult.wge(c(0.9), c(0.8))
```

```{r}
# first the first 5 weights (l = 6)
weights <- psi.weights.wge(c(0.9, 0.8), lag.max = 5)
# calculate the sum of the squared weights
sum(c(1,weights)^2)
```


## Assignment 3

```{r}
mse <- function(y.true, y.pred){
  mean( (y.true - y.pred)^2 )
}

# load the data
data <- read.csv('./AmtrakPassengersMonthly.csv')

test.set <- data$Ridership[(159-11):159]

# model 1
m1.phi = c(0.5511, 0.1680, -0.0145, 0.0651, 0.1388,
           -0.2966, 0.1539, 0.1270, -0.1815, 0.0364,
           0.1456, 0.6287, -0.3832, -0.0199, -0.1679) 

m1.fcast <- fore.arma.wge(data$Ridership, phi = m1.phi, n.ahead = 12, lastn = T)

m1.mse <- mse(m1.fcast$f, test.set)

print('The MSE for model 1 is')
print(m1.mse)

# model 2
m2.phi = c(-0.02709541,  0.74213105)
m2.theta = c(-0.5844596,  0.3836931)

m2.fcast <- fore.arma.wge(data$Ridership,
                          phi = m2.phi,
                          theta = m2.theta,
                          n.ahead = 12,
                          lastn = T)

m2.mse <- mse(m3.fcast$f, test.set)

print('The MSE for model 2 is')
print(m2.mse)

# model 3
m3.phi = 0.306943
m3.theta = 0.7431719

m3.fcast <- fore.arma.wge(data$Ridership,
                          phi = m3.phi,
                          theta = m3.theta,
                          n.ahead = 12,
                          lastn = T)

m3.mse <- mse(m2.fcast$f, test.set)

print('The MSE for model 3 is')
print(m3.mse)
```






























