---
title: "Live_Session_Assignemnt_6"
author: "Stuart Miller"
date: "February 8, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tswge)

# read in data
data1 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTest.csv')
data2 <- read.csv('./daily-climate-time-series-data/DailyDelhiClimateTrain.csv')
data <- rbind(data2, data1)

data['seq'] <- seq_along(data[, 'humidity'])
```

## Assignment 1

Looking at your time series from the first day of class, you addressed its stationarity before. Does either a signal plus noise, ARIMA, or ARUMA (seasonal) model seem appropriate?

Since the sample autocorrelations are slowly damped, it would expect that an ARIMA model would work.

```{r}
plotts.sample.wge(data$humidity)
```

## Assignment 2

Copy and paste the factor table for a seasonal model with s = 7.

```{r}
factor.wge(c(rep(0,6),1))
```


## Assignemnt 3

Idetify the type of model for the following models

### Part A

$$
\left( 1 - B^4 \right)X_t = \left( 1 + 0.3 B\right) \left( 1 - 0.6B + 0.8 B^2 \right)
$$

This appears to be a seasonal (quarterly) model with an MA(3) noise component.

```{r}
# generate a series from this model
theta <- mult.wge(c(-0.3), c(0.6, -0.8))
x <- gen.aruma.wge(n = 200, theta = theta$model.coef, s = 4, plot = F)
# remove the seasonal component
x_diff <- artrans.wge(x, phi.tr = c(0,0,0,1))

# plot both series
plotts.sample.wge(x)
plotts.sample.wge(x_diff)

# fit models to the noise
aic5.wge(x_diff)
```

```{r}
factor.wge(c(0,0,0,1))
```


### Part B

$$
(1+.3B) X_t = \left(1-.6B+.8B^2\right) a_t
$$

This is an ARMA(1,2) model

```{r}
# generate a series from this model
theta <- mult.wge(c(-0.3), c(0.6, -0.8))
x <- gen.aruma.wge(n = 200, phi = c(-0.3), theta = c(0.6, -0.8), plot = F)

# plot series
plotts.sample.wge(x)
# fit models to the noise
aic5.wge(x)
```

```{r}
factor.wge(c(0.6, -0.8))
factor.wge(c(-0.3))
```


### Part C

$$
\left(1-.1B-.99B^2 +.013B^3 +.2078B^4 +.0888B^5 +.00864B^6\right)(1-B)^2 \left(1-B^{12} \right) X_t = (1-.6B+.8B^2) a_t
$$

This is a full ARUMA model with difference order of 1 and monly seasonality.
The noise structure is ARMA(6,2)

```{r}
# generate a series from this model
x <- gen.arima.wge(n = 200, 
                   phi = c(.1,.99,-.013,-.2078,-.0888,-.00864), 
                   theta = c(0.6, -0.8),
                   d = 2,
                   s = 12,
                   plot = F)
# remove the seasonal component
x_diff <- artrans.wge(x, phi.tr = c(seq(0,11),1))
x_diff <- artrans.wge(artrans.wge(x_diff, phi.tr = 1), phi.tr = 1)

# plot both series
plotts.sample.wge(x)
plotts.sample.wge(x_diff)

# fit models to the noise
aic5.wge(x_diff,  p = 0:8, q = 0:5)
```


## Assignment 4

Pick a stock and download the last year's worth of daily stock price data.

* Plot the data. 
* Take the first difference, and then estimate the structure of differenced data.  
* From this information, suggest a model for the data. 

```{r}
# load in stock data
stock <- read_csv('./TXN.csv')
```

```{r}
# plot the data
plotts.sample.wge(stock$Close)
```

```{r}
# take the first difference
x <- artrans.wge(stock$Close, phi.tr = 1)
aic5.wge(x)
```



















































