---
title: "homework_4"
author: "Stuart Miller"
date: "February 3, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(tidyverse)
```

# Problem 3.2

Plot the following for the two models listed below

1. A realization
2. The theoretical autocorrelations
3. The theoretical spectral density

Then find the theoretical mean and varaince (assuming a white noise variance of 1) and the mean and varaince of the realization.

$$
X_t - 0.95 X_{t-1} = a_t
$$

Generate a realization and plot the true autocorrelations and spectral density.

```{r}
x <- plotts.true.wge(phi = 0.95)
```

This AR model is in 0 mean form.
Since the expectation value of the mean of an AR process is the mean, the theoretical mean is 0. 
The theoretical variance is given by 

$$
\sigma_X = \frac{\sigma_a^2}{1 - \phi_1^2}
$$

Then the theoretical varaince is

```{r}
1 / (1 - 0.95^2)
```

The mean and variance of the realization are

```{r}
var(x$data)
mean(x$data)
```


$$
(X_t - 10) - 1.5 (X_{t-1} - 10) +0.9 (X_t - 10) = a_t
$$

Generate a zero-mean realization and plot the true autocorrelations and spectral density.

```{r}
x <- plotts.true.wge(phi = c(1.5, -0.9))
```

Plot the realization with the correct mean offset.

```{r}
x10 <- x$data+10
plotts.wge(x$data+10)
```

Since the expectation value of the mean of an AR process is the mean, the theoretical mean is 10. 
The theoretical variance is given by 

$$
\sigma_X = \frac{\sigma_a^2}{1 - \phi_1 \rho_1 - \phi_1 \rho_2}
$$

Then the theoretical varaince is

```{r}
1 / (1 - 1.5 * x$aut1[2] + 0.9 * x$aut1[3])
```

The mean and variance of the realization are

```{r}
var(x10)
mean(x10)
```


# Problem 3.3

Sections a, b, and d only.

Determine if the following models are stationary. Explain the answer.

$$
X_t - 1.55 X_{t-1} + X_{t-2} - 0.25 X_{t-3} = a_t
$$

This model is stationary because all of the roots are outside the unit circle.

```{r}
factor.wge(phi = c(1.55, -1, 0.25))
```



$$
X_t -2 X_{t-1} + 1.76 X_{t-2} - 1.6 X_{t-1} + 0.77 X_{t-4} = a_t
$$

Since this AR process has a root inthe unit circle, it is not stationary.

```{r}
factor.wge(phi = c(2, -1.76, 1.6, -0.77))
```


$$
X_t - 1.9 X_{t-1} + 2.3 X_{t-2} - 2 X_{t-1} + 1.2 X_{t-4} -0.4 X_{t-5} = a_t
$$

Since all the roots are outside the unit circle, this model is stationary.

```{r}
factor.wge(c(1.9, -2.3, 2, -1.2, 0.4))
```






















