---
title: "homwork_5"
author: "Stuart Miller"
date: "February 3, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

# Problem 3.1

Plot the following for the two models listed below

1. A realization
2. The theoretical autocorrelations
3. The theoretical spectral density

Then find the theoretical mean and varaince (assuming a white noise variance of 1) and the mean and varaince of the realization.

$$
X_t - 25 = a_t - 0.95 a_{t-1} + 0.9 a_{t-2} - 0.855a_{a-3}
$$

Generate a realization and plot the true autocorrelations and spectral density.

```{r}
x <- plotts.true.wge(theta = c(0.95, -.9, .855))
x25 = x$data + 25
```

Replot the realization with the correct mean

```{r}
plotts.wge(x25)
```

The the mean is 25 and the theoretical variance is given by

$$
\sigma^2 = \sigma_a^2(1 + \theta_1^2 + \theta_2^2 + \theta_3^2)
$$

Assuming the white noise variance is 1, we have

```{r}
1 + 0.95^2 + 0.9^2 + 0.855^2
```

The mean and variance are calculated as follows

```{r}
mean(x25)
var(x25)
```

# Problem 3.3

All moving average models are stationary regardless of were the roots are located.

# Problem 3.6

For the following model

* find the order of the model
* write out the equation in operator form
* determine if the model is stationary and invertable

$$
X_t - X_{t-1} + 0.26 X_{t-2} +0.64 X_{t-3} -0.576 X_{t-4} = a_t - 2.4a_{t-1} + 2.18 a_{t-2} - 0.72 a_{t-3}
$$

Factor the equation on both sides

```{r}
factor.wge(c(1, -0.26, -0.64, 0.576))
```

```{r}
factor.wge(c(2.4, -2.18, 0.72))
```


Based on the factoring above, the factor $(1-0.8B)$ will cancel out.
So this is an ARMA(3,2) model,

This model can be expressed as 

$$
(1-B+0.9B^2)(1+0.8B) X_t = (1-1.6B+0.9B^2) a_t
$$

The factor tables show that all roots are outside of the unit circle.
Thus, the model is stationary and invertable.


# Problem 3.7

## Part A

The factor table for the model

```{r}
factor.wge(c(0.1, -0.5, -.08, 0.24))
```

Since all the roots of the AR process are outside the unit circle, it is stationary.
Since this is an AR model it is invertable.

The true spectral density and autocorrelations

```{r}
x <- plotts.true.wge(phi = c(0.1, -0.5, -.08, 0.24))
```

The roots indicate a dominate peak at $f=0.237$ and some presence at $f=0$ and $f=0.5$.
There is some evidence in the period of oscillation (4) in the autocorrelations.

## Part B

AR factor table

```{r}
factor.wge(c(1.3, -0.4))
```

MA factor table

```{r}
factor.wge(c(1.9))
```

This process is stationary because all of the roots of the AR process are outside the unit circle.
However, the MA component contains a root inside the unit circle.
Thus the process is not invertable.

## Part C

The AR factor table

```{r}
factor.wge(c(1.9))
```

MA factor table

```{r}
factor.wge(c(1.3, -0.4))
```

The process is not stationary because the AR component has a root inside the unit circle.
However, the process is invertable because the roots of the MA components are outside the unit circle.

## Part D

The AR factor table

```{r}
factor.wge(c(2.95, -3.87, 2.82, -0.92))
```

MA factor table

```{r}
factor.wge(c(0.9))
```

The process is both stationary and invertable because all the roots are outside the unit circle.


The true spectral density and autocorrelations

```{r}
x <- plotts.true.wge(phi = c(2.95, -3.87, 2.82, -0.92), theta = c(0.9))
```

## Part E

The AR factor table

```{r}
factor.wge(c(1, 0.49, -0.9, 0.369))
```

MA factor table

```{r}
factor.wge(c(-1, -1, -0.75))
```

This model is stationary and invertable because all the roots of the characteristic equations are outside of the unit circle.

The AR process has dominate roots at 0 and 0.5.
We expect to see peaks in the spectral density at these points.
The MA process has a dominate root at 0.2623.
We would expect to see a minimum in the spectral density at this point.

```{r}
x <- plotts.true.wge(phi = c(1, 0.49, -0.9, 0.369), theta = c(-1, -1, -0.75))
```


# Problem 3.8

## Part A

The following MA process is not invertable because it has roots inside the unit circle.

```{r}
factor.wge(c(0.5, 5))
```

```{r}
true.arma.aut.wge(theta = c(0.5, 5), lag.max = 2, plot = F)
```

```{r}
mult.wge(.4, -0.5)
```

An invertable MA process with the same autocorrelations is shown below.

```{r}
factor.wge(c(-0.1, 0.2))
```


```{r}
true.arma.aut.wge(theta = c(-0.1, 0.2), lag.max = 2, plot = F)
```

## Part B

The following MA model is not invertable because the roots are inside the unit circle.

```{r}
factor.wge(c(2, -1.5))
```

```{r}
true.arma.aut.wge(theta = c(2, -1.5), lag.max = 2, plot = F)
```

An invertable MA process with the same autocorrelations is shown below.

```{r}
factor.wge(c(-1.3333, -0.66667))
```


```{r}
true.arma.aut.wge(theta = c(-1.33333, -0.66667), lag.max = 2, plot = F)
```


# Problem 3.11

AR Factor Table

```{r}
factor.wge(c(1, 0.49, -0.9, 0.369))
```


MA Factor Table

```{r}
factor.wge(c(-1, -1, -0.75))
```

Based on the factor tables above, this is a stationary, invertable process

The first 10 $\psi$s are shown below.

```{r}
psi.weights.wge(phi = c(1, 0.49, -0.9, 0.369), theta = c(-1, -1, -0.75), lag.max = 10)
```


# Problem 3.13i

```{r}
factor.wge(c(2.2, -2.1, 0.8))
```

Based on the factor table, fit an AR(3) and examine two additive components.
As expected the dominate behavior comes from the pair of complex roots.
The AR(1) provides minimal impact on the realization.

```{r}
X = gen.arma.wge(n = 100, phi = c(2.2, -2.1, 0.8), plot = F, sn = 42)
factor.comp.wge(X, p = 3, ncomp = 2)
```


























