---
title: "homework_3"
author: "Stuart Miller"
date: "January 19, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(tidyverse)
```

# Problem 2.1

Create the signal and store it in `X_t`.

```{r}
X_t = gen.sigplusnoise.wge(n = 100, coef = c(3, 1.5), freq = c(0.05, 0.35), psi = c(0, 2))
plotts.sample.wge(X_t)
```

For the following plot the realization, sample autocorrelations, and parzen window.
Discuss the cyclic behavior.

a. Apply a third order low-pass Butterworth filter with cut off at 0.2
b. Apply a third order high-pass Butterworth filter with cut off at 0.2
c. Apply a third order high-pass Butterworth filter with cut off at 0.2 and a third order low-pass Butterworth filter with cut off at 0.2

```{r}
X_t.lowpass <- butterworth.wge(X_t, order = 3, cutoff = 0.2, type = "low")
X_t.highpass <- butterworth.wge(X_t, order = 3, cutoff = 0.2, type = "high")
X_t.bandpass <- butterworth.wge(X_t.lowpass$x.filt, order = 3, cutoff = 0.2, type = "high")
```

## Part A

Compared to the original data,
the low-pass filter as removed the high frequency (the oscillations characterized by frequencies above 0.2).
The realization appears to be a low freuqnecy signal.

```{r}
plotts.sample.wge(X_t.lowpass$x.filt)
```

## Part B

Compared to the original data,
the high-pass filter as removed the low frequency (the oscillations characterized by frequencies low 0.2).
The realization appears to be a high freuqnecy signal.

```{r}
plotts.sample.wge(X_t.highpass$x.filt)
```

## Part C

Compared to the original data,
the band-pass filter as removed most of the oscillation (the oscillations characterized by frequencies low and above 0.2).
The realization appears to be relatively flat with some noise.

```{r}
plotts.sample.wge(X_t.bandpass$x.filt)
```

# Problem 2.3

Apply a 1st order difference filter to the time series.

Like the butterworth high-pass filter, the differencing filter suppresses the low frequency. 
However, the supression from the difference filter is not as strong - 
there is still evidence of a frequency at 0.05.
Note that a 20(n) dB/decade roll off.

```{r}
X_t.diff.filter <- artrans.wge(X_t, 1)
plotts.sample.wge(X_t.diff.filter)
```

# Additional Problem

Apply a 5-point moving average to the series you created in 2.1.
How does it compare to the difference and Butterworth filters?
Specifically, is it a low pass or high pass filter?

The 5-point moving average filter is a low-pass filter.
It does not reject the high frequency as strongly as the low-pass butterworth filter.
However, based on the realization, the moving average filter may be practically sufficient for filtering the signal.

```{r}
ma.5p <- stats::filter(X_t, rep(1,5)/5)
plotts.sample.wge(na.omit(ma.5p))
```

# Problem 3.4

Generate an AR(1) process with n = 100, phi = +/- 0.9, +/- 0.5, and sigma^2_a = 1.

a. plot the true autocorrelations  
b. plot the true spectral density

```{r}
plotts.true.wge(n = 200, phi = 0.9)
plotts.true.wge(n = 200, phi = -0.9)
plotts.true.wge(n = 200, phi = 0.5)
plotts.true.wge(n = 200, phi = -0.5)
```

c. calculate sigma^2_X

$$
\sigma^2_X = \frac{\sigma^2_a}{1 - \phi^2} 
$$

From the equation above, the answers are as follows:

* phi with abs value 0.9: 0.81
* phi with abs value 0.5: 1.33

d. and e. plot the realizations and sample autocorrelations

```{r}
xp5 <- gen.arma.wge(n = 200, phi = 0.5, sn = 1)
xm5 <- gen.arma.wge(n = 200, phi = -0.5, sn = 1)
xp9 <- gen.arma.wge(n = 200, phi = 0.9, sn = 1)
xm9 <- gen.arma.wge(n = 200, phi = -0.9, sn = 1)

plotts.sample.wge(xp5)
plotts.sample.wge(xm5)
plotts.sample.wge(xp9)
plotts.sample.wge(xm9)
```
























