---
title: "Homework 11"
author: "Stuart Miller"
date: "March 17, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(tswgewrapped)
```

# Homework 11

## Problem 9.1

### Part A

Fit a model to a realization generated from the known model using the best ID based on AIC.

```{r}
phi = c(1.5, -1.21, 0.455)
gen <- gen.arma.wge(200, phi = phi, sn = 100)
aic5.wge(gen, p = 0:8, q = 0:3)
vals <- est.arma.wge(gen, p = 6, q = 2)
# check for whiteness of the residuals
white_noise_eval(vals$res, 6, 2)
```

Plot the ARMA spectral density and the estimated windowed spectral density.

The spectral density plots are similar.
However, the is a strange notch in the true spectral density plot.

```{r}
a <- parzen.wge(gen)
a <- plotts.true.wge( phi = vals$phi, theta = vals$theta)
```


Validate the model based on a forecast.

The last 10 data points will be used for validation.
We can see that the forecast follows the validation steps very well.

```{r}
fore.arma.wge(gen, phi = vals$phi, theta = vals$theta,
              lastn = T, n.ahead = 10)
```

Generate some simulations from the fitted model to compare to the original realization

Realizations appear to be similar.
In general, they tend to express the type of oscillatory behavior seen in the original realization.

```{r}
gen.arma.wge(200, phi = vals$phi, theta = vals$theta)
```


### Part B

Fit a model to a realization generated from the known model using the best ID based on AIC.

The model selected by AIC appears to reduce the residuals to white noise.

```{r}
phi = c(1.6, -1.7, 1.28, -0.72)
factor.wge(phi)
gen <- gen.arma.wge(200, phi = phi, sn = 100)
aic5.wge(gen, p = 0:8, q = 0:3)
vals <- est.arma.wge(gen, p = 4, q = 0)
# check for whiteness of the residuals
white_noise_eval(vals$res, 4, 0)
```

Plot the ARMA spectral density and the estimated windowed spectral density.

The spectral density plots are similar.
The spectral density of the fitted model captures the primary shape of the sample spectral density.

```{r}
a <- parzen.wge(gen)
a <- plotts.true.wge( phi = vals$phi, theta = vals$theta)
```


Validate the model based on a forecast.

The last 10 data points will be used for validation.
We can see that the forecast follows the validation steps very well.

```{r}
fore.arma.wge(gen, phi = vals$phi, theta = vals$theta,
              lastn = T, n.ahead = 10)
```

Generate some simulations from the fitted model to compare to the original realization

Realizations appear to be similar.
In general, they tend to express the type of oscillatory behavior seen in the original realization.

```{r}
gen.arma.wge(200, phi = vals$phi, theta = vals$theta)
```


### Part C

Fit a model to a realization generated from the known model using the best ID based on AIC.

The model selected by AIC appears to reduce the residuals to white noise.

```{r}
phi = c(1 , -1.76, 0.8, -0.72)
factor.wge(phi)
gen <- gen.arma.wge(200, phi = phi, sn = 100)
aic5.wge(gen, p = 0:8, q = 0:3)
vals <- est.arma.wge(gen, p = 4, q = 0)
# check for whiteness of the residuals
white_noise_eval(vals$res, 4, 0)
```

Plot the ARMA spectral density and the estimated windowed spectral density.

The spectral density plots are similar.
The spectral density of the fitted model captures the primary shape of the sample spectral density.

```{r}
a <- parzen.wge(gen)
a <- plotts.true.wge( phi = vals$phi, theta = vals$theta)
```


Validate the model based on a forecast.

The last 10 data points will be used for validation.
We can see that the forecast follows the validation steps very well.

```{r}
fore.arma.wge(gen, phi = vals$phi, theta = vals$theta,
              lastn = T, n.ahead = 10)
```

Generate some simulations from the fitted model to compare to the original realization

Realizations appear to be similar.
In general, they tend to express the type of oscillatory behavior seen in the original realization.

```{r}
gen.arma.wge(200, phi = vals$phi, theta = vals$theta)
```



## Problem 9.2

### Part A

The residuals of the fit model do not appear to be consistent with white noise, 
but the ACF does appear to be consistent.

```{r}
data("wtcrude")

plotts.sample.wge(wtcrude)

aic5.wge(wtcrude, p = 0:8, q = 0:2)
vals <- est.arma.wge(wtcrude, p = 7, q = 2)
# check for whiteness of the residuals
white_noise_eval(vals$res, 7, 2)
```

Plot the ARMA spectral density and the estimated windowed spectral density.

The spectral density plots are similar.
The spectral density of the fitted model captures the primary shape of the sample spectral density.

```{r}
a <- parzen.wge(wtcrude)
a <- plotts.true.wge( phi = vals$phi, theta = vals$theta)
```


Validate the model based on a forecast.

The last 10 data points will be used for validation.
We can see that the the movement of the realization is captures by the forecast limits,
but the forcast is a bit lower than the actual behavior.

```{r}
fore.arma.wge(wtcrude, phi = vals$phi, theta = vals$theta,
              lastn = T, n.ahead = 10)
```

Generate some simulations from the fitted model to compare to the original realization

Realizations appear to be similar.

```{r}
gen.arma.wge(200, phi = vals$phi, theta = vals$theta)
```






### Part B

The residuals of the fitted model do not appear to be consistent with white noise.
The ACF ACF appears to be consistent with white noise and the Ljung-Box tests fail to reject the null hypothesis.

```{r}
data("eco.mort30")

plotts.sample.wge(eco.mort30)

aic5.wge(eco.mort30, p = 0:8, q = 0:3)
vals <- est.arma.wge(eco.mort30, p = 6, q = 2)
# check for whiteness of the residuals
white_noise_eval(vals$res, 6, 2)
```

Plot the ARMA spectral density and the estimated windowed spectral density.

The spectral density plots are similar.
The spectral density of the fitted model captures the primary shape of the sample spectral density.

```{r}
a <- parzen.wge(eco.mort30)
a <- plotts.true.wge( phi = vals$phi, theta = vals$theta)
```


Validate the model based on a forecast.

The last 50 data points will be used for validation.
The forecast does not appear to track the movements of the realization.
The realization moves down, but the forecase is up to the mean.

```{r}
fore.arma.wge(eco.mort30, phi = vals$phi, theta = vals$theta,
              lastn = T, n.ahead = 50)
```

Generate some simulations from the fitted model to compare to the original realization

Realizations appear to be similar.
In general, they tend to express the type of oscillatory behavior seen in the original realization.

```{r}
gen.arma.wge(200, phi = vals$phi, theta = vals$theta)
```




### Part C

The residuals appear to be consistent with white noise. 
Additionally, the ACF appear to be consistent with white noise and the Ljung-Box tests fail to reject the null hypothesis.

```{r}
data("hadley")

plotts.sample.wge(hadley)

aic5.wge(hadley, p = 0:8, q = 0:3)
vals <- est.arma.wge(hadley, p = 4, q = 2)
# check for whiteness of the residuals
white_noise_eval(vals$res, 4, 2)
```

Plot the ARMA spectral density and the estimated windowed spectral density.

The spectral density plots are similar.
The spectral density of the fitted model captures the primary shape of the sample spectral density.

```{r}
a <- parzen.wge(hadley)
a <- plotts.true.wge( phi = vals$phi, theta = vals$theta)
```


Validate the model based on a forecast.

The last 20 data points will be used for validation.
We can see that the forecast does not forecast the movement of the realization very well.

```{r}
fore.arma.wge(hadley, phi = vals$phi, theta = vals$theta,
              lastn = T, n.ahead = 20)
```

Generate some simulations from the fitted model to compare to the original realization

Realizations appear to be similar.

```{r}
gen.arma.wge(200, phi = vals$phi, theta = vals$theta)
```


## Problem 9.3

```{r}
gen <- gen.aruma.wge(200, s = 6, phi = 0.8, sn = 100) - 25
est <- est.ar.wge(gen, p = 10, type = 'burg')
transformed <- artrans.wge(gen, phi.tr = c(rep(0,5),1))
aic5.wge(transformed)
est <- est.arma.wge(transformed, p = 1)
```

Check For White Noise Residuals

Based on the plots and the Ljung-Box tests, the residuals appear to be consistent with white noise.

```{r}
# check for whiteness of the residuals
white_noise_eval(est$res, 1, 0)
```

Plot the ARMA spectral density and the estimated windowed spectral density.

The spectral density plots are similar.
The spectral density of the fitted model captures the primary shape of the sample spectral density.

```{r}
a <- parzen.wge(transformed)
a <- plotts.true.wge( phi = est$phi, theta = est$theta)
```


Validate the model based on a forecast.

The last 20 data points will be used for validation.
We can see that the forecast does not forecast the movement of the realization very well.

```{r}
fore.aruma.wge(transformed, phi = est$phi, theta = est$theta, s = 6,
              lastn = T, n.ahead = 20)
```

Generate some simulations from the fitted model to compare to the original realization

Realizations appear to be similar.

```{r}
gen.aruma.wge(200, phi = vals$phi, theta = vals$theta, s = 6)
```



