---
title: "homework_6"
author: "Stuart Miller"
date: "February 9, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
data("hadley")
```

# Homework 6

## Problem 5.1

### Part A

Plot the data.

```{r}
plotts.wge(hadley)
```


### Part B

Fit a regression model as `hadley` ~ `time`.

```{r}
time <- seq_along(hadley)
model <- lm(hadley ~ time)
summary(model)
```

### Part C

Find and plot the residuals for the linear model.



```{r}
plot(model$residuals)
```


### Part D

The residuals could have come from a stationary model.
Constant mean apepars to be reasonable.
The varaince appears to be constant.
The autocorrelation for the first half of the data and the second half of the data are very similar.

```{r}
resid.size <- length(model$residuals)
acf(model$residuals[1:resid.size/2])
acf(model$residuals[resid.size/2:resid.size])
```


### Part E

The autocorrelations of `hadley` are more slowly damping than the residuals.
There is some suggestion of oscillatory behavior, but it is not strong.

The autocorrelations of the residuals of the linear model fall off much more rapidly.
Additionally, there is stronger evidence of oscillator behavior.

```{r}
plotts.sample.wge(hadley)
```

```{r}
plotts.sample.wge(model$residuals)
```


## Problem 5.2

### Part A

Generate 3 realizations for the following model.
Do any of the models appear to have a linear trend?

The first model appears to have a negative linear trend.

$$
X_t - 0.66 X_{t-1} + 0.02 X_{t-2} - 0.10 X_{t-3} - 0.24 X_{t-4} = a_t
$$

```{r}
phi <- c(0.66, -0.02, 0.1, 0.24)
x1 <- gen.arma.wge(n = 150, phi = phi, plot = T, sn = 1)
x2 <- gen.arma.wge(n = 150, phi = phi, plot = T, sn = 42)
x3 <- gen.arma.wge(n = 150, phi = phi, plot = T, sn = 35)
```

### Part B

Plot 4 realizations for the following model and compare to the hadley realization.

The forth realization looks similar to the hadley realization.
The trends appear to be strong in the other realizations.

$$
X_t = -0.526 + 0.0044t + Z_t
$$

Where

$$
(1 - 0.614 B  + 0.044 B^2 - 0.077 B^3 - 0.0206 B^4)  Z_t = 0
$$

```{r}
phi <- c(0.614, -0.044, 0.077, 0.0206)
x1 <- gen.sigplusnoise.wge(n = 150, b0 = -0.526, b1 = 0.0044,
                           phi = phi, sn = 1, vara = 0.01)
x2 <- gen.sigplusnoise.wge(n = 150, b0 = -0.526, b1 = 0.0044,
                           phi = phi, sn = 42, vara = 0.01)
x3 <- gen.sigplusnoise.wge(n = 150, b0 = -0.526, b1 = 0.0044,
                           phi = phi, sn = 37, vara = 0.01)
x4 <- gen.sigplusnoise.wge(n = 150, b0 = -0.526, b1 = 0.0044,
                           phi = phi, sn = 35, vara = 0.01)
```


## Problem 5.3 (a, b, d)

For the following models generate a realziation and discuss the observable featrues.
Additionalyl identify if the model is

* Stationary
* Non-stationary
* Seasonal (and non-stationary)

$$
\left(1- 1.2B + 0.8 B^2 \right) (1-B) X_t = a_t
$$

$$
\left(1- 1.2B + 0.8 B^2 \right) (1-B)^2 X_t = a_t
$$

$$
\left(1- 1.2B + 0.8 B^2 \right) (1-B)^2 \left( 1 - B^4 \right)
\left(1 + B + B^2 \right) X_t = a_t
$$

### Realizations

```{r}
phi = c(1.2, -0.8)
gen.aruma.wge(n = 200, phi = phi, d = 1, sn = 3)
gen.aruma.wge(n = 200, phi = phi, d = 2, sn = 3)
md <- gen.aruma.wge(n = 200, phi = phi, d = 1, s= 4, lambda = c(-1,-1), sn = 3)
```

### Part A

Since this model has a root on the unit circle, it is non-stationary.
The realization exhibits a psuedo-periodic behavior,
which is consistent with the complex root in the AR porition of the model.

```{r}
factor.wge(phi = phi)
```

### Part B

The realization almost appears to be a linear trend.
This model has two roots on the unit circle, which dominate the realiztion behavior.

### Part D

This model contains a quarterly seasonal factor, 2 roots on the unit circle,
and addtional non-stationary factors.
The model is dominated by the roots on the unit circle. 
The quarterly oscillatory behavior is also expressed.

```{r}
model.d <- mult.wge(phi,c(1,1),c(rep(0,3),1),c(-1,-1))
factor.wge(model.d$model.coef)
```

```{r}
plotts.sample.wge(md)
```


## Problem 5.5

### Part A

The factored model is 

$$
(1-B)^2 (1 + B + B^2)(1 - B + 0.5 B^2) = 1-1.7B+0.8B^2
$$

There are no canceling factors, the model is an ARUMA(2,4,2).

#### AR Factor Table

```{r}
factor.wge(c(3, -4.5, 5, -4, 2, -0.5))
```

#### MA Factor Table

```{r}
factor.wge(c(1.7,-0.8))
```


### Part B

The factored model is 

$$
(1-B) (1 + B + 0.8B^2) (1 + 0.5 B^2) (1 -0.5B) = 0
$$

There are no canceling factors, the model is an ARUMA(5,1,0).

#### AR Factor Table

```{r}
factor.wge(c(0.5, -0.3, 0.95, -0.3, 0.35, -0.2))
```

#### MA Factor Table

There are no MA factors


### Part C

The factored model is 

$$
(1+B) (1-1.5B+1B^2) (1-1B+0.8B^2) (1+0.5B^2) = (1+0.9B^2)
$$

There are no canceling factors, the model is an ARUMA(4,3,2).

#### AR Factor Table

```{r}
factor.wge(c(1.5, -1.3, -0.35, 1, - 1.35 , 0.7, -0.4))
```

#### MA Factor Table

```{r}
factor.wge(c(0, -0.9))
```


### Part D

The factored model is 

$$
(1-B)^2 (1+B) (1+B+B^2) (1+0.5B) = (1+0.9B) (1-0.9B)
$$

There are no canceling factors, the model is an ARUMA(1,5,2).

#### AR Factor Table

```{r}
factor.wge(c(0.5, 0.5, 0, 1, -0.5, -0.5))
```

#### MA Factor Table

```{r}
factor.wge(c(0, 0.81))
```


## Problem 5.7

### Part A

The roots of unit are

* $\pm 1$
* $-0.5000 \pm 0.8660i$
* $0.5000 \pm 0.8660i$

```{r}
factor.wge(c(rep(0,5),1))
```

### Part B and Part D

There is clear evidence of psuedo-periodic behavior.
The autocorrelations do not match what would be expected from a theoretical prespective, but the strongest autocorrelations occur at intervals of 6, which is the expected behavior.
The frequencies shown in the Parzen are similar to the ones in the factor table.

```{r}
x <- gen.aruma.wge(n = 120, s = 6, sn = 6)
plotts.sample.wge(x)
```

### Part C

The true autocorrelations have a magnitude of 1 and occur periodically in intervals of 6.
There are frequencies shown in true spectral density match the frequencies shown in the factor table.

```{r}
plotts.true.wge(phi = c(rep(0,5),0.9999))
```

### Part D














































































