---
title: "homework_1"
author: "Stuart Miller"
date: "January 5, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

# Homework 1

## Problem 1.1

Done by hand; not transcribed

## Problem 1.2

Load the west Texas crude oil dataset and the Pennsylvania monthly temperatures dataset from tswge.
Plot the asmple autocorrelations, periodograms, and a Parzen window-based spectrial density estimate.
Describe the behavior of the time series.

### West Texas Crude

The line plot of the realization shows that the data exhibits a non-cyclic behavior.
The autocorrelation plot shows that there is a strong correlation between samples that are nearby (gamma > 0.5 for k <= 13).
The Parzen window and Periodogram show a peak at 0 with no indication of cyclic behavior.
There is no indication of the drop after the peak of x_t = 100.

```{r}
# load the west texas data
data(wtcrude)
# make the plots
plotts.sample.wge(wtcrude)
```

### Pennsylvania Month Temperature

The data exhibits cyclic behavior with a peak at about 0.08 on the periodogram and Parzen window.
We would expect to see a cyclic behavior over the year (lambda = 1/12 = 0.0.83).
The autocorrelations appear to be slowly damped with a period of 12.

```{r}
# load the west penn temperature data
data(patemp)
# make the plots
plotts.sample.wge(patemp)
```

## Problem 1.4

Realization 1
  * autocorrelation: the time series exhibits wandering behavior, which corresponds to strong positive autocorrelations starting from 0 (d).
  * spectral density: the time series exhibits wandering behavior, which corresponds a peak at 0 and exponentially approaching 0 as the frequency increases (a).
  
  Realization 2
  * autocorrelation: the time series appears to have little structure, which corresponds to a no autocorrelations (a)
  * spectral density: the time series appears to have little structure, which corresponds to a flat spectrum line (d)

Realization 3
  * autocorrelation: the time series exhibits cyclic behavior with a period of about 10, which matches the periodic autocorrelations in (b).
  * spectral density: the time series exhibits cyclic behavior with a period of about 10, which matches the spectral density with a peak near 0.1 (c)

Realization 4
  * autocorrelation: the time series exhibits cyclic behavior with more than one frequency, matching autocorrelation plot (c), which shows osciallations in the main periodic correlations
  * spectral density: the time series exhibits cyclic behavior with more than one frequency, matching the spectral density plot (b), which has two peaks

## Problem 1.6

Generate a realization of the following equation of length 100.
Discuss the plots that show the periodic behavior.

$$
X_t = 3cos(2\pi(0.05)t) + 1.5cos(2\pi(0.35)t) + 2) + N(0,1)
$$

The sample autocorrelation plot shows the dominant frequency well with the period of 20 visible (0.05).
The Parzen window and the periodogram both suggest a frequeny at 0.05 and 0.35.

```{r}
X_t = gen.sigplusnoise.wge(n = 100, coef = c(3, 1.5), freq = c(0.05, 0.35), psi = c(0, 2))
plotts.sample.wge(X_t)
```




