---
title: "homework_9"
author: "Stuart Miller"
date: "March 14, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

# Homework 9

## Problem 7.3

### Part A

Find the model estimates for an AR(2) and AR(8) fitted to the sunspot data with mle, YW, and burg methods.

```{r}
data("sunspot.classic")

# fit the AR(2)
cat('ML estimate of AR(2)\n')
est.ml.ar2 <- est.ar.wge(sunspot.classic, p = 2)
cat('YW estimate of AR(2)\n')
est.yw.ar2 <- est.ar.wge(sunspot.classic, p = 2, type = 'yw')
cat('Burg estimate of AR(2)\n')
est.b.ar2 <- est.ar.wge(sunspot.classic, p = 2, type = 'burg')


# fit the AR(2)
cat('ML estimate of AR(2)\n')
est.ml.ar8 <- est.ar.wge(sunspot.classic, p = 8)
cat('YW estimate of AR(2)\n')
est.yw.ar8 <- est.ar.wge(sunspot.classic, p = 8, type = 'yw')
cat('Burg estimate of AR(2)\n')
est.b.ar8 <- est.ar.wge(sunspot.classic, p = 8, type = 'burg')
```

### Part B

Find the model estimates for an AR(1) fitted to the given data with mle, YW, and burg methods.

```{r}
dat <- c(26.8, 27.8, 30.0, 31.6, 33.0,
         34.2, 34.1, 33.2, 31.7, 31.5,
         31.9, 31.5, 30.0, 29.4, 29.9)

cat('ML estimate of AR(1)\n')
est.ml.ar2 <- est.ar.wge(dat, p = 1)
cat('YW estimate of AR(1)\n')
est.yw.ar2 <- est.ar.wge(dat, p = 1, type = 'yw')
cat('Burg estimate of AR(1)\n')
est.b.ar2 <- est.ar.wge(dat, p = 1, type = 'burg')
```


### Part C

Find the model estimates for an AR(2) fitted to the given data with mle, YW, and burg methods.

```{r}

cat('ML estimate of AR(2)\n')
est.ml.ar2 <- est.ar.wge(dat, p = 2)
cat('YW estimate of AR(2)\n')
est.yw.ar2 <- est.ar.wge(dat, p = 2, type = 'yw')
cat('Burg estimate of AR(2)\n')
est.b.ar2 <- est.ar.wge(dat, p = 2, type = 'burg')
```


## Problem 7.4

Generate a realization from each model and find the mle, YW, and burg for each.

Burg and ML estimates appear to be pretty good,
YW is not that good on higher order models and root close to unity.

```{r}
phi1 <- c(1.3, -0.6)
phi2 <- c(3.1, -4.1, 2.63, -0.72)
phi3 <- c(2.69, -2.583, 0.891)
```

### Part A

```{r}

dat <- gen.arma.wge(100, phi = phi1)

cat('ML estimate of AR(2)\n')
est.ml.ar <- est.ar.wge(dat, p = 2)
cat('YW estimate of AR(2)\n')
est.yw.ar <- est.ar.wge(dat, p = 2, type = 'yw')
cat('Burg estimate of AR(2)\n')
est.b.ar <- est.ar.wge(dat, p = 2, type = 'burg')
```


### Part B

```{r}

dat <- gen.arma.wge(100, phi = phi2)

cat('ML estimate of AR(2)\n')
est.ml.ar <- est.ar.wge(dat, p = 4)
cat('YW estimate of AR(2)\n')
est.yw.ar <- est.ar.wge(dat, p = 4, type = 'yw')
cat('Burg estimate of AR(2)\n')
est.b.ar <- est.ar.wge(dat, p = 4, type = 'burg')
```

### Part C

```{r}

dat <- gen.arma.wge(100, phi = phi3)

cat('ML estimate of AR(2)\n')
est.ml.ar <- est.ar.wge(dat, p = 3)
cat('YW estimate of AR(2)\n')
est.yw.ar <- est.ar.wge(dat, p = 3, type = 'yw')
cat('Burg estimate of AR(2)\n')
est.b.ar <- est.ar.wge(dat, p = 3, type = 'burg')
```




## Problem 7.5

```{r}
parzen.wge(sunspot.classic)
plotts.true.wge(n = length(sunspot.classic), p = est.ml.ar2$phi)
plotts.true.wge(n = length(sunspot.classic), p = est.ml.ar8$phi)
```

## Problem 7.6

```{r}
dat <- gen.arma.wge(100, phi = phi2)

cat('ML estimate of AR(2)\n')
est.ml.ar <- est.ar.wge(dat, p = 4)
cat('YW estimate of AR(2)\n')
est.yw.ar <- est.ar.wge(dat, p = 4, type = 'yw')
cat('Burg estimate of AR(2)\n')
est.b.ar <- est.ar.wge(dat, p = 4, type = 'burg')

parzen.wge(dat)
plotts.true.wge(length(dat), p = est.ml.ar$phi)
plotts.true.wge(length(dat), p = est.yw.ar$phi)
plotts.true.wge(length(dat), p = est.b.ar$phi)
```


































































