---
title: "homework 10"
author: "Stuart Miller"
date: "March 14, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

# Homework 10

## Problem 8.1

Are the following datasets consistent with white noise?

The realizations and autocorrelations of 1a and 1d appear to consistent with white noise.

The ACF of 1b does not appear to be consistent with white noise.

The realization of 1c does not apprear to be consistent with white noise.

```{r}
data("prob8.1a")
data("prob8.1b")
data("prob8.1c")
data("prob8.1d")

plotts.sample.wge(prob8.1b)
```


## Problem 8.2

### Part A

Only one root of the fit model is close to the roots of the actual model.

```{r}
phi = c(1.5, -1.21, 0.455)
factor.wge(phi)
gen <- gen.arma.wge(200, phi = phi, sn = 100)
aic5.wge(gen, p = 0:8, q = 0:3)
vals <- est.arma.wge(gen, p = 6, q = 2)
mean(gen-50)
vals$avar
```

**Model specification**

$$
\left( 1 + 0.299B - 0.228B^2 + 0.109B^3 + 0.606B^4 - 0.306B^5 - 0.0979B^6 \right) 
\left( X_t - 50.107 \right)
 = a_t
\,\,\,\,
\sigma_a^2 = 0.8052
$$

### Part B

The fit appears to capture the original model fairly well.

```{r}
phi = c(1.6, -1.7, 1.28, -0.72)
factor.wge(phi)
gen <- gen.arma.wge(200, phi = phi, sn = 100)
aic5.wge(gen, p = 0:8, q = 0:3)
vals <- est.arma.wge(gen, p = 4, q = 0)
mean(gen-200)
vals$avar
```

**Model specification**

$$
\left( 1 - 1.5810B + 1.7833B^2 - 1.3939B^3 + 0.7949B^4  \right)  
\left( X_t -200.02 \right)
= a_t
\,\,\,\,
\sigma_a^2 = 0.8125
$$


### Part C

The fit appears to capture the original model fairly well.

```{r}
phi = c(1 , -1.76, 0.8, -0.72)
factor.wge(phi)
gen <- gen.arma.wge(200, phi = phi, sn = 100)
aic5.wge(gen, p = 0:8, q = 0:3)
vals <- est.arma.wge(gen, p = 4, q = 0)
mean(gen-200)
vals$avar
```

**Model specification**

$$
\left( 1 - 1.0121B + 1.8126B^2 - 0.8880B^3 + 0.7861B^4  \right)  
\left( X_t -199.99 \right)
= a_t
\,\,\,\,
\sigma_a^2 = 0.843
$$

## Problem 8.3

### Part A

```{r}
data("wtcrude")

plotts.sample.wge(wtcrude)

aic5.wge(wtcrude, p = 0:5, q = 0:2)
est <-  est.arma.wge(wtcrude, 3, 2)
factor.wge(est$theta)
mean(wtcrude)
est$avar
```

$$
\left(1 - 2.3254B + 2.0124B^2 - 0.6694B^3  \right) \left( X_t  + 21.98 \right) = 
\left( 1 - 1.0174B + 0.6010B^2  \right) a_t 
\,\,\,\,
\sigma_a^2 = 21.9835
$$


### Part B

```{r}
data("eco.mort30")

plotts.sample.wge(eco.mort30)

aic5.wge(eco.mort30, p = 0:8, q = 0:3)
est <-  est.arma.wge(eco.mort30, 6, 2)
factor.wge(est$theta)
mean(eco.mort30)
est$avar
```

$$
\left(1 - 0.5538B + 0.2136B^2 - 0.8615B^3 
+ 0.4133B^4 - 0.3464B^5 + 0.1528B^6   \right) 
\left( X_t  + 8.97 \right) = 
\left( 1 + 1.0228 B + 0.8847 B^2 \right) a_t 
\,\,\,\,
\sigma_a^2 = 0.064
$$

### Part C

```{r}
data("hadley")

plotts.sample.wge(hadley)

aic5.wge(hadley, p = 0:8, q = 0:3)
est <-  est.arma.wge(eco.mort30, 2, 2)
factor.wge(est$theta)
mean(hadley)
est$avar
```

$$
\left(1 - 0.5255B - 0.4578B^2 \right) 
\left( X_t  - 0.168 \right) = 
\left( 1 + 1.0676 B + 0.3428 B^2 \right) a_t 
\,\,\,\,
\sigma_a^2 = 0.0669
$$



## Problem 8.5

The following are the non-stationary factors from the realization

Factor                 Roots                Abs Recip    System Freq 
1+0.9992B             -1.0008               0.9992       0.5000
1+1.0101B+0.9907B^2   -0.5098+-0.8657i      0.9954       0.3347
1-1.0089B+0.9876B^2    0.5108+-0.8670i      0.9938       0.1653




```{r}
gen <- gen.aruma.wge(200, s = 6, phi = 0.8, sn = 100) - 25
est.ar.wge(gen, p = 10, type = 'burg')
transformed <- artrans.wge(gen, phi.tr = c(rep(0,5),1))
aic5.wge(transformed)
est.arma.wge(transformed, p = 1)
```


## Problem 8.6

### Part A

```{r}
data("freight")
plotts.sample.wge(freight)
diff1 <- artrans.wge(freight, phi.tr = 1)
diff1.12 <- artrans.wge(freight, phi.tr = c(rep(0,11),1))
aic5.wge(diff1.12)
est.ar.wge(diff1.12, p = 2)
```

### Part B

```{r}
data("freeze")
plotts.sample.wge(freeze)
est.ar.wge(freeze, p = 15)
diff1 <- artrans.wge(freeze, phi.tr = c(1.97, -1))
plotts.sample.wge(diff1)
aic5.wge(diff1)
est <- est.arma.wge(diff1, q = 3)
factor.wge(est$theta)

plotts.sample.wge( est$res )

```

### Part C

```{r}
data("lavon")
plotts.sample.wge(lavon)

est.ar.wge(lavon, 15)


diff1 <- artrans.wge(lavon, phi.tr = c(rep(0,3),1))

est.ar.wge(diff1, 15)

diff1.s4 <- artrans.wge(diff1, phi.tr = c(rep(0,3),1))

aic5.wge(diff1, p = 0:17)
est.arma.wge(diff1, p = 2, q = 1)
```




























































