---
title: "ARIMA and ARUMA Models"
author: "Stuart Miller"
date: "February 8, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

# Signal + Noise Generation

## Linear Signals

```{r}
n = 100

# generate linear signals + noise
gen.sigplusnoise.wge(n, b0 = 2, b1 = 4, vara = 100)
# this one is just noise
gen.sigplusnoise.wge(n, b0 = 0, b1 = 0, vara = 10)
# this is a linear signal with an AR(1)
gen.sigplusnoise.wge(n, b0 = 0, b1 = 0, phi = 0.975, vara = 10)
```

## Periodic Signal

$$
X_t = 5 cos (2 \pi (0.1) t + 2.5) +Z_t | Z_t ~ AR(1)
$$

```{r}
gen.sigplusnoise.wge(n, coef = c(5,0), freq = c(0.1,0), psi = c(0.25,0), phi = 0.975, vara = 4)
```

## AR(4) Model

$$
\left( 1 - 0.975B \right) \left( 1-0.2B + 0.45B^2 \right) \left( 1 + 0.53 \right) X_t = a_t
$$


```{r}
params = mult.wge(c(0.975), c(0.2,-0.45), c(-0.53))
params$model.coef
gen.arma.wge(n = 160, phi = params$model.coef, vara = 1)
```


## ARIMA(0,1,0)

Realization is strickly wandering with root at 1.

```{r}
x = gen.arima.wge(200, phi = 0, var = 1, d = 1, sn  = 31)
```

Sample autocorrelations are slowly damped

```{r}
acf(x)
```


# Differencing Data

The following command can be used to make differences in data

```{r}
# let x be a series
x = c(1,3,6,10,25)
# compute the first difference y
y <- artrans.wge(x, phi.tr = 1)
```

# ARIMA Analysis

Get a factor table of a model in factored form

```{r}
phi = mult.wge(c(0.6,-0.8),c(1), c(1))
factor.wge(phi$model.coef)
```

```{r}
x = gen.arima.wge(n = 500, phi = c(0.6, -0.8), d = 2, theta = -0.3, sn = 35)
plotts.sample.wge(x)
```


```{r}
y = artrans.wge(artrans.wge(x, phi.tr = 1), phi.tr = 1)
aic5.wge(y)
```


```{r}
parzen.wge(y)
```


# Generating ARUMA and Seasonal Data

Generate data for the following seasonal model.

$$
\left( 1 - B^4 \right) X_t = a_t
$$


```{r}
x1 = gen.aruma.wge(n = 80, s = 4, sn = 6)
plotts.sample.wge(x1)
```


Generate data for s4 with ARMA(2,1)

```{r}
x2 = gen.aruma.wge(n = 80, phi = c(1, -0.6), s = 4, theta = -0.5, sn = 6)
plotts.sample.wge(x2)
```

```{r}
factor.wge(c(1,-0.6))
factor.wge(-0.5)
```


## Stationarize ARUMA

Generate data from the following model,
then remove the seasonal component to stationarize the data.

$$
\left( 1 - 0.6 B + 0.8 B^2 \right) \left( 1 -B^12 \right) X_t = \left(1 -0.3 B + 0.7 B^2 \right) a_t
$$

```{r}
# generate the data
x = gen.aruma.wge(n = 500, phi = c(0.6, -0.8), theta = c(0.3, -0.7), s = 12, sn = 37)
x = artrans.wge(x, phi.tr = c(rep(0,11),1))
plotts.sample.wge(x)
aic5.wge(x)
```


## Factoring Seasonal Models

Check for sim

```{r}
factor.wge(c(-0.3, 1.2, 0.4, 0, 0.5, rep(0,6), -1, 0.3, -1.2, -0.4))
```

```{r}
factor.wge(c(rep(0,11),1))
```












