---
title: "Moving Average"
author: "Stuart Miller"
date: "January 29, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(tidyverse)
```

# Generating Data and Plots

## generating MA(1) data

```{r}
gen.arma.wge(n = 100, theta = -0.99)
gen.arma.wge(n = 100, theta = 0.99)
gen.arma.wge(n = 100, theta = -0.99, sn = 5)
gen.arma.wge(n = 100, theta = 0.99, sn = 5)
```



## True plots

```{r}
plotts.true.wge(theta = c(0.99))
```


## Generating MA(2) Data

```{r}
gen.arma.wge(n = 100, theta = c(0.9, -0.4))
```


## True plots

```{r}
plotts.true.wge(theta = c(0.9, -0.4))
```

# Invertability

`factor.wge` can be used to factor MA models.
This can be used to determine if models have roots in the unit circle.

This model would be considered invertable.

```{r}
factor.wge(phi = c(1.6, -0.9))
```

This model is not invertable because one of the roots is in the unit circle.

```{r}
factor.wge(phi = c(1.6, 0.9))
```

# Blending MA and AR to Make ARMA

## AR factors

```{r}
plotts.true.wge(phi = c(0.3, 0.9, 0.1, -0.8075))
factor.wge(phi = c(0.3, 0.9, 0.1, -0.8075))
```

## MA factors

```{r}
plotts.true.wge(theta =  c(-0.9, -0.8, -0.71))
factor.wge(c(-0.9, -0.8, -0.71))
```

## ARMA

```{r}
plotts.true.wge(theta =  c(-0.9, -0.8, -0.71), phi = c(0.3, 0.9, 0.1, -0.8075))
```

# Model section by AIC

```{r}
SWA <- read_csv('./swadelay.csv')
plotts.wge(SWA$weather_delay)
```


```{r}
aic5.wge(SWA$weather_delay)
```

# Calculating psi-weights

## MA(1)

```{r}
psi.weights.wge(theta = 0.8, lag.max = 5)
```

## AR(1)

```{r}
psi.weights.wge(phi = 0.8, lag.max = 5)
```
















