---
title: "Parameter_Estimation"
author: "Stuart Miller"
date: "2/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

# Model Estimation



## Maximum Likelihood Estimation

```{r}
x <- gen.arma.wge(n = 200,
                  phi = c(0.3, -0.7), 
                  theta = -0.4,
                  vara = 4,
                  sn = 27)
est.arma.wge(x,2,1)
```

## Burg Estimation

```{r}
x <- gen.arma.wge(n = 200,
                  phi = c(0.3, -0.7), 
                  vara = 4,
                  sn = 27)
est.ar.wge(x,2,1, type = 'burg')
```

# Model Identification

## Step 1 -  Check for White Noise

The following appears to be white noise based on the realization and autocorrelations.

```{r}
mwn <- read.csv('./maybewhitenoise1.csv')$x
plotts.sample.wge(mwn)
```

The following appears to have autocorrelations consistent with white noise,
but the realization does not appear to be consistent with white nouse.

```{r}
mwn <- read.csv('./maybewhitenoise2.csv')$x
plotts.sample.wge(mwn)
```

## Step 2 - Select p, q

This can be done with AIC or a similar criterion.

$$
AIC = ln \left( \hat{\sigma}^2 \right) + 2 \left( \frac{  p + q + 1  }{n} \right)
$$

* BIC imposes a stronger penalty for adding model parameters than AIC
* AICc is a middle groud between AIC and BIC

```{r}
inflation <- read.csv('./inflation.csv')
aic5.wge(inflation$Inflation, type = 'bic')
```


# PACF

```{r}
data <- read.csv('./armawhatpq1.csv')$x
acf(data)
```


```{r}
aic5.wge(data)
aic5.wge(data, type = 'bic')
```


```{r}
pacf(inflation$Inflation)
```



# Putting it all together

## Example 1

Plots show that this does not look like a white noise process.

```{r}
data <- read.csv('./putittogether.csv')$x
plotts.sample.wge(data)
```

The best fit for this is an ARMA(p, q) based on AIC.

```{r}
aic5.wge(data)
```

Get the model parameter estimates.

```{r}
est.arma.wge(data, p = 2, q = 2)
```



## Example 2

Plots show that this does not look like a white noise process.

```{r}
data <- read.csv('./texasgasprice.csv')$Price
plotts.sample.wge(data)
```

The best fit for this is an ARMA(p, q) based on AIC.

```{r}
aic5.wge(data)
```

Get the model parameter estimates.

```{r}
est <- est.arma.wge(data, p = 2)
```

```{r}
fore.arma.wge(data, phi = est$phi, n.ahead = 8)
```




























