---
title: "forecasting"
author: "Stuart Miller"
date: "February 11, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

## AR Forecasting

### AR(1) Example

```{r}
data("fig6.1nf")
plotts.wge(fig6.1nf)
```

```{r}
fore.arma.wge(fig6.1nf, phi = 0.8, n.ahead = 20, plot = T, limits = F)
```

### AR(2) Example

```{r}
x2 <- gen.arma.wge(n = 75, phi = c(1.6, -0.8), sn = 24)
x2 <- x2 + 25
fore.arma.wge(x2, phi = c(1.6, -0.8), n.ahead = 50)
```


## ARMA Forecasting

AR(2,1) example.

```{r}
x1 <- gen.arma.wge(n = 75, phi = c(1.6, -0.8), theta = -0.9, sn = 24)
fore.arma.wge(x1, phi = c(1.6, -0.8), theta = -0.9, n.ahead = 20)
```





## Checking Forecasts

Forecast on the Lynx data and calculate MSE.

```{r}
data("llynx")
plotts.wge(llynx)
l_AR4 = fore.arma.wge(llynx, phi = c(1.3, -0.7, 0.1, -0.2),
                      n.ahead = 30, limits = F, lastn = T)
MSE= mean((l_AR4$f - llynx[85:114])^2)
MSE

```


```{r}
swd <- read.csv('./swadelay.csv')

AR12 <- fore.arma.wge(swd$arr_delay, 
                      phi = c(0.44, 0.02, -0.12, 0.08, 0, 0.02,
                              0.06, -0.09, 0.06, 0.07, 0.02, 0.37),
                     n.ahead = 30, lastn = T, limits = F)

AR12.mse = mean((AR12$f - swd$arr_delay[148:177])^2)

print(AR12.mse)

ARMA12 <- fore.arma.wge(swd$arr_delay,
                        phi = c(0.34, 0.07, -0.11, 0.07, 0.01,
                                0.02, 0.06, -0.09, 0.05, 0.07,
                                0.02, 0.39), 
                        theta = -0.12,
                        n.ahead = 30, lastn = T, limits = F)

ARMA12.mse = mean((ARMA12$f - swd$arr_delay[148:177])^2)

print(ARMA12.mse)
```


## ARIMA Forecasting

Forecasting on an ARIMA(0,1,0)

$$
(1-B) \left( X_t - \mu \right) = a_t
$$

```{r}
xd = gen.aruma.wge(n = 75, d = 1, sn = 74)
fore.aruma.wge(xd, d = 1, n.ahead = 5, limits = F)
```

### Other Forecast Examples

```{r}
xd = gen.aruma.wge(n = 50, phi = 0.8, d = 1, sn = 15)
fore.aruma.wge(xd, d = 1, n.ahead = 5, limits = F)
```

```{r}
fore.aruma.wge(xd, phi = 0.8, d = 1, n.ahead = 5, limits = F)
```


```{r}
fore.aruma.wge(xd, d = 2, n.ahead = 5, limits = F)
```

## ARIMA with Seasonality Forecasting

Forecast from the following model

$$
\left( 1- B^4 \right) X_t = a_t
$$

```{r}
x = gen.aruma.wge(n = 20, s = 4, sn = 6)
fore.aruma.wge(x, s = 4, n.ahead = 8, lastn = F, plot = T)
```

```{r}
x = gen.aruma.wge(n = 20, s = 4, sn = 6)
fore.aruma.wge(x, s = 4, n.ahead = 8, lastn = T, plot = T)
```

### Forecasting on the Airline Data

```{r}
data("airlog")
f = fore.aruma.wge(airlog, phi = c(-0.36, -0.05, -0.14, -0.11, 0.04,
                               0.09, -0.02, 0.02, 0.17, 0.03,
                               -0.1, -0.38),
               d= 1, s = 12,
               n.ahead = 36,
               lastn = T)

mean((f$f - airlog[(144-35):144])^2)
```


## Signal + Noise Forecasting

Generate a linear signal with an AR(2) noise structure

```{r}
x <- gen.sigplusnoise.wge(n = 50, b0 = 10, b1 = 0.2, phi = c(0.8, -0.6))
```


```{r}
fore.sigplusnoise.wge(x, n.ahead = 20)
```




### Amazon Prediction

```{r}
amazon <- read.csv('./AMZN.csv')
```


```{r}
f <- fore.sigplusnoise.wge(amazon$Adj.Close, n.ahead = 30)
```







































































