---
title: "mlr_correlated_errors"
author: "Stuart Miller"
date: "March 18, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(vars)
library(tidyverse)
library(RColorBrewer)
```

## Variable Lags

Consider the attached data set in WhatIsTheLag.csv. 
Use the cross-correlation function (ccf()) to find evidence of the lag in which X1 is related to Y. 
That is, which k has the most evidence of Xt-k being related to Yt?

```{r}
dat <- read_csv('./whatisthelag.csv')
ccf(dat$Y, dat$X1_1, lag.max = 10)
```


## VAR Example

```{r}
x1.25=c( -1.03, 0.11, -0.18, 0.20, -0.99, -1.63, 1.07, 2.26, -0.49, -1.54, 0.45, 0.92,
        -0.05, -1.18, 0.90, 1.17, 0.31, 1.19, 0.27, -0.09, 0.23, -1.91, 0.46, 3.61, -0.03)
x2.25=c( -0.82, 0.54, 1.13, -0.24, -0.77, 0.22, 0.46, -0.03, -0.59, 0.45, 0.59, 0.15,
        0.60, 0.13, -0.04, 0.12, -0.96, 0.23, 1.81, -0.01, -0.95, -0.55, -0.15, 0.71, 0.90)
x1=x1.25[1:20]
x2=x2.25[1:20] 
p1=aic.wge(x1,p=0:8,q=0:0)
# aic picks p=2
x1.est=est.ar.wge(x1,p=p1$p)
fore.arma.wge(x1,phi=x1.est$phi,n.ahead=5,lastn=FALSE,limits=FALSE)
p2=aic.wge(x2,p=0:8,q=0:0)
# aic picks p=2
x2.est=est.ar.wge(x1,p=p2$p)
fore.arma.wge(x2,phi=x2.est$phi,n.ahead=5,lastn=FALSE,limits=FALSE)
#
# VAR and VARselect are from CRAN package vars
X=cbind(x1,x2)
VARselect(X, lag.max = 6, type = "const",season = NULL, exogen = NULL)
#VARselect picks p=5 (using AIC)
lsfit=VAR(X,p=5,type="const")
preds=predict(lsfit,n.ahead=5)
# preds$fcst$x1[1,1] - [5,1] are the VAR forecasts for x1. Similar for x2
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues")) # Change color pallet to make distinguishable.

summary(lsfit)
```

## In Class VAR

```{r}
dat <- read_csv('./businesssales.csv')[2:5]
dat
VARselect(dat, lag.max = 6, type = "const",season = NULL, exogen = NULL)
lsfit=VAR(dat,p=2,type="const")
summary(lsfit)
AIC(lsfit)
```

## Forecast LA_Cmort_Study

Your goal is to utilize all given information to provide the most useful forecasts for the next 20 weeks of cardiac mortality. 

```{r}
# get the data
dat <- read_csv('./la_cmort_study.csv')[2:4]

# create a data split
end <- length(dat$cmort)
val.start <- end - 20 + 1
val.set <- dat[val.start : end, ]
dat.train <- dat[1 : (val.start - 1), ]

```

```{r}
plotts.sample.wge(dat$cmort)
plotts.sample.wge(dat$temp)
plotts.sample.wge(dat$part)
```

### Cross-Correlation Plots

```{r}
ccf(dat$cmort, dat$temp)
ccf(dat$cmort, dat$part)
```

### Select, Fit, and Validate VAR Models

```{r}
# VAR analysis
# fit with trend
VARselect(dat.train, lag.max = 40, type = "trend",season = NULL, exogen = NULL)
lsfit.trend.17 <- VAR(dat.train,p=17,type="trend")
summary(lsfit.trend.17)
AIC(lsfit.trend.17)
# get forecast on next 20
preds.trend.17 <- predict(lsfit.trend.17,n.ahead=20)

mean((preds.trend.17$fcst$cmort[ , 1] - val.set$cmort)^2)

# fit without trend
VARselect(dat.train, lag.max = 40, type = "const",season = NULL, exogen = NULL)
lsfit.const.14=VAR(dat.train,p=14,type="const")
summary(lsfit.const.14)
AIC(lsfit.const.14)
# get forecast on next 20
preds.const.14 <- predict(lsfit.const.14,n.ahead=20)
preds.const.14$fcst$cmort[ , 1]

mean((preds.const.14$fcst$cmort[ , 1] - val.set$cmort)^2)

# generate plot
plot(seq_along(dat.train$cmort), dat.train$cmort, xlim = c(350, 510))
points(seq(val.start, end), val.set$cmort)
points(seq(val.start, end), preds.const.14$fcst$cmort[ , 1], type = 'b', pch = 15)

plot(seq_along(dat.train$cmort), dat.train$cmort, xlim = c(350, 510))
points(seq(val.start, end), val.set$cmort)
points(seq(val.start, end), preds.trend.17$fcst$cmort[ , 1], type = 'b', pch = 15)

```

### Forecast Future

```{r}
lsfit <- VAR(dat,p=17,type="trend")
preds <- predict(lsfit,n.ahead=20)

plot(seq_along(dat$cmort), dat$cmort, xlim = c(350, 530))

points(seq((end + 1), (end + 20)),
       preds$fcst$cmort[ , 1], type = 'b', pch = 15)

fanchart(preds, names = c('cmort'), colors = brewer.pal(n = 8, name = "Blues"), xlim = c(350, 530))
```

## Calculate Error on Sales Forecasts

```{r}
dat <- read_csv('./businesssales.csv')[2:5]

end <- length(dat$sales)

# fit on the data before the prediction length
ksfit <- lm(sales ~ ad_tv + ad_online + discount, data = dat[1:(end - 5), ])
aic.wge(ksfit$residuals, 0:8, q = 0)
fit <- arima(dat$sales[1:(end - 5)], order = c(7, 0, 0), xreg = dat[1:(end - 5) , 2:4])
# predict on the last 5 values
preds <- predict(fit, n.ahead = 5, newxreg = dat[ (end - 4):end, 2:4])

# show the predictions
preds$pred

```


```{r}
mean( (preds$pred - dat$sales[ (end - 4):end])^2 )
```

```{r}
plot(seq_along(preds$pred), preds$pred, ylim = c(45, 65))
points(seq_along(preds$pred), dat$sales[ (end - 4):end], pch = 15)
```



```{r}
VARselect(dat[1:(end - 5), ], lag.max = 6, type = "const",season = NULL, exogen = NULL)
lsfit <- VAR(dat[1:(end - 5), ], p=2, type="const")
summary(lsfit)
AIC(lsfit)
```

```{r}
preds <- predict(lsfit, n.ahead = 5)
preds$fcst$sales[ , 1]

mean( (preds$fcst$sales[ , 1] - dat$sales[ (end - 4):end])^2 )
```


```{r}
plot(seq_along(preds$fcst$sales[ , 1]), preds$fcst$sales[ , 1], ylim = c(45, 65))
points(seq_along(preds$fcst$sales[ , 1]), dat$sales[ (end - 4):end], pch = 15)
```











































































