---
title: "Model_Building"
author: "Stuart Miller"
date: "March 12, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

# Model Building

Example model

$$
(1 - 1.6 B + 0.9 B^2 ) (X_t-10) = (1 -0.8 B)a_t
$$


```{r}
x = gen.arma.wge(n = 100, phi = c(1.6, -0.9), theta =0.8, sn = 67) + 10
plotts.sample.wge(x)
aic5.wge(x)
x21 <- est.arma.wge(x, 2, 1)
x21$phi
x21$theta
x21$avar
mean(x)

```


$$
(1 - 1.62 B + 0.91 B^2 ) (X_t-10.08) = (1 -0.87 B)a_t \,\,\,\, \sigma^2_a = 1.08
$$

Check whitened residuals

```{r}
plotts.sample.wge(x21$res)
ljung.wge(x21$res,p=2,q=1)
ljung.wge(x21$res,p=2,q=1, K = 48)
```

## Sunspot Data

```{r}
data("sunspot.classic")
aic5.wge(sunspot.classic, p = 0:10)
plotts.sample.wge(sunspot.classic)
```

## The Most Recent Sunspot Data

```{r}
data <- read.csv('./SN_y_tot_V2.0.csv', sep = ';', header = F)
plotts.sample.wge(data$V2)
```


```{r}
shift <- length(data$V2) / 2
end <- length(data$V2)

acf(data$V2[1 : shift])
acf(data$V2[shift : end])
```

Used BIC becuase AIC tends to pick a model with a large number of parameters

Selected model orders

* p = 4
* q = 4

```{r}
aic5.wge(data$V2, p = 0:15, q = 0:5, type = 'aic')
aic5.wge(data$V2, p = 0:15, q = 0:5, type = 'bic')
```

```{r}
model.arma <- est.arma.wge(data$V2, 9, 0)
model.arma.f <- fore.arma.wge(data$V2,
                              phi = model.arma$phi,
                              theta = model.arma$theta,
                              n.ahead = 15,
                              lastn = T)
ljung.wge(model.arma$res, p = 9, q = 0, K = 48)
```

```{r}
artrans.wge(data$V2,
            phi.tr = model.arma$phi)
```


```{r}
start <- length(data$V2) - 15 + 1
val.set <- data$V2[start : end]

mean( ( model.arma.f$f - val.set )^2 )
```

```{r}
model.arma <- est.arma.wge(data$V2, 24, 0)
```


```{r}
factor.wge(c(rep(0,10),1))
```



```{r}
diff.11 <- artrans.wge(data$V2, phi.tr = c(rep(0,10),1))
aic5.wge(diff.10, p = 0:15, type = 'aic')
aic5.wge(diff.10, p = 0:15, type = 'bic')
```

```{r}
model.sarma <- est.arma.wge(diff.11, 13)
model.sarma.f <- fore.aruma.wge(data$V2,
                                 phi = model.sarma$phi,
                                 theta = model.sarma$theta,
                                 d = 0,
                                 s = 10, 
                                 lastn = T,
                                 n.ahead = 15)

plotts.sample.wge(model.sarma$res)

ljung.wge(model.sarma$res, p = 13, K = 48)
```



```{r}
mean( ( model.sarma.f$f - val.set )^2 )
```



```{r}
fore.aruma.wge(data$V2,
               phi = model.arma$phi,
               theta = model.arma$theta,
               d = 0,
               s = 0, 
               lastn = F,
               n.ahead = 10)
```


## The Accuspike Webhits Data

From the sample plots, the data appears to be consistent with an AR(3).
There is a prominate frequency near 1.4 and a prominate frequency near 0.
However, the frequency near 1.4 could be a weekly seasonal pattern since the data is taken daily.
The realization appears to be stationary except for the last few points.

```{r}
data <- read.csv('./Accuspike.csv', header = T)
realization <- data$Active.Users[1:180]
plotts.sample.wge(realization)
```

```{r}
end <- length(realization) -1
shift <- end / 2

acf(realization[ 1 : shift ])
acf(realization[ shift : end])

```


Removing a seasonal component of s = 7 appears to reduce the autocorrelations.

```{r}
diff.s7 <- artrans.wge(realization, phi.tr = c(rep(0,6),1))

end <- length(diff.s7) -1
shift <- end / 2
acf(diff.s7[ 1 : shift ])
acf(diff.s7[ shift : (end)])
```

```{r}
plotts.sample.wge(diff.s7)
vals <- est.ar.wge(diff.s7, 24, type = 'burg')
```


```{r}
aic5.wge(diff.s7, p = 0:10, q = 0:3)
aic5.wge(diff.s7, p = 0:10, q = 0:3, type = 'bic')
```



```{r}
est <- est.arma.wge(diff.s7, p = 8, q = 0)
est$avar
mean(realization)
```

$$
\left(1 -0.36B +0.0002B^2 - 0.08B^3 - 0.071B^4 + 0.17B^5  + 0.0015B^6 + 0.45B^7 - 0.3275B^8 \right) 
\left( 1-B^7 \right)
\left(X_t - 83.85 \right)
= a_t \,\,\,\, \sigma_a^2 = 303.77
$$


```{r}
diff.s7.ar8 <- artrans.wge(diff.s7, phi.tr = est$phi)
plotts.sample.wge(diff.s7.ar8)
ljung.wge(diff.s7.ar8, p = 8, K = 24)
```


```{r}
sim <- gen.aruma.wge(200, phi = est$phi, s = 7, theta = est$theta)
parzen.wge(sim)
```


```{r}
fore.aruma.wge(realization,
               s = 7,
               phi = est$phi,
               n.ahead = 7*3,
               lastn = F)
```








































