---
title: "nn_mlp"
author: "Stuart Miller"
date: "March 23, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nnfor)
library(tseries)
library(tswge)
dat <- read_csv('./swadelay.csv')
```

## Nerual Networks and MLPs

### Univariate MLPs

```{r}
SWA.train <- ts(dat$arr_delay[1:141],
                start = c(2004, 1),
                frequency = 12)
SWA.test <- ts(dat$arr_delay[142:177],
                start = c(2015, 10),
                frequency = 12)

set.seed(2)

fit.mlp <- mlp(SWA.train, reps = 50, comb = 'mean')
fit.mlp
```

Represetation of model.

Pink nodes are for seasonal inputs.
Gray nodes are the lags

```{r}
plot(fit.mlp)
```



```{r}
fore.mlp <- forecast(fit.mlp, h = 36)
plot(fore.mlp)
```



```{r}
ASE <- mean( (SWA.test - fore.mlp$mean)^2 )
ASE
```


```{r}
set.seed(2)
fit.mlp <- mlp(SWA.train, reps = 20, comb = 'median',
               allow.det.season = F,
               lags = c(1,2,3,4,5,6,7,8,9,10,11,12))
fit.mlp
fore.mlp <- forecast(fit.mlp, h = 36)
ASE <- mean( (SWA.test - fore.mlp$mean)^2 )
ASE
```



```{r}
set.seed(2)
fit.mlp <- mlp(SWA.train, reps = 20, comb = 'median',
               allow.det.season = F,
               difforder = c(12))
fit.mlp
fore.mlp <- forecast(fit.mlp, h = 36)
ASE <- mean( (SWA.test - fore.mlp$mean)^2 )
ASE
```



```{r}
set.seed(2)
fit.mlp <- mlp(SWA.train, difforder = c(1,6,12),
              allow.det.season = FALSE, reps = 100)
fit.mlp
fore.mlp <- forecast(fit.mlp, h = 36)
ASE <- mean( (SWA.test - fore.mlp$mean)^2 )
ASE
```


#### Airline Data

```{r}
data("airlog")
set.seed(2)
liar.train <- ts(airlog[1:108],
                 frequency = 12,
                 start = c(1949, 1))
liar.test <- ts(airlog[109:144],
                frequency = 12,
                start = c(1958, 1))

fit.mlp <- mlp(liar.train,
               hd.auto.type = 'cv')
fit.mlp
```


```{r}
fore.mlp <- forecast(fit.mlp, h = 36)
ASE <- mean( (liar.test - fore.mlp$mean)^2 )
ASE
```

### Multivariate MLP

```{r}
dat <- read_csv('./businesssales.csv')[2:5]

```


```{r}
set.seed(2)
tBS80 <- ts(dat$sales[1:80])
tBSx <- data.frame(ad_tv = ts(dat$ad_tv),
                   ad_online = ts(dat$ad_online, frequency = 7),
                   discount <- ts(dat$discount))
fit <- mlp(tBS80, xreg = tBSx,
               hd.auto.type = 'cv')
f <- forecast(fit, h = 20, xreg = tBSx)
plot(dat$sales[81:100], type = 'l')
lines(seq(1,20), f$mean, col = 'blue')

ASE <- mean( (dat$sales[81:100] - f$mean)^2 )
ASE
```


```{r}
fit
```


### For Live Session

#### Melanoma Subspot Data

Model the sunspot/melanoma data with a MLP. 
Use the model to provide eight-step-ahead forecasts.
Finally, calculate the ASE, and compare your NN model to the VAR model we estimated previously.  

```{r}
data <- read.csv('./Melanoma_Sunspot.csv')
plotts.sample.wge(data$Melanoma)
plotts.sample.wge(data$Sunspot)
```



```{r}
set.seed(2)

end <- length(data$Melanoma)
shift <- 8

dat.test <- ts(data$Melanoma[(end - shift + 1) : end],
               start = 30)
dat.train <- ts(data$Melanoma[1: (end - shift)],
                start = 1)

x.reg <- data.frame(Sunspot = ts(data$Sunspot, frequency = 10))

fit <- mlp(dat.train,
           xreg = x.reg,
           reps = 50)

f <- forecast(fit,
              h = 8, 
              xreg = x.reg)

plot(f)

mean( (dat.test - f$mean)^2 )

plot(seq_along(dat.test) + length(dat.train), dat.test, type = 'l', 
     main = 'Neural Net Forecast', xlab = 'Time', ylab = 'Melanoma')
lines(f$mean, col = 'blue')


```

### LA Study

 Model the Schumway pollution data with a MLP model, and provide 10-step-ahead forecasts.
 Calculate the ASE, and add your analysis to a slide in your Unit 13 For Live Session PowerPoint deck.  

```{r}
data <- read.csv('./la_cmort_study.csv')
end <- length(data$cmort)

test <- ts(data$cmort[(end - 10 + 1) : end], 
           start = end - 10 + 1)
train <- ts(data$cmort[1 : (end - 10)])

x.reg <- data.frame(
  part = ts(data$part),
  temp = ts(data$temp)
  )

# get seasonal factor 
val <- plotts.sample.wge(train)
s <- round(1 / mean(c(val$freq[9], val$freq[10])))
s

set.seed(2)

fit <- mlp(train, 
           xreg = x.reg,
           reps = 50)

f <- forecast(fit,
              h = 10, 
              xreg = x.reg)

plot(f)

mean( (test - f$mean)^2 )

plot(seq_along(test) + length(train), test, type = 'l', 
     main = 'Neural Net Forecast', xlab = 'Time', ylab = 'Cmort')
lines(f$mean, col = 'blue')
```


#### Custom Series

 Model your time series with a MLP, 
 and compare it with ARMA/ARIMA models you have fit to it in the past. 

```{r}
# read in data
data1 <- read.csv('./live_session_assignments/daily-climate-time-series-data/DailyDelhiClimateTest.csv')
data2 <- read.csv('./live_session_assignments/daily-climate-time-series-data/DailyDelhiClimateTrain.csv')

set.seed(2)

plotts.sample.wge(data2$humidity)

test <- ts(data1$humidity, start = 1463)
train <- ts(data2$humidity, start = 1)

fit <- mlp(train, difforder = c(375))

tswgewrapped::evaluate_residuals(fit$fitted - train)

f <- forecast(fit, h = length(test))

ASE <- mean( (test - f$mean)^2 )
ASE

plot(seq_along(test) + length(train), test, type = 'l', 
     main = 'Neural Net Forecast', xlab = 'Time', ylab = 'Humidity')
lines(f$mean, col = 'blue')
```

```{r}
diff <- artrans.wge(train, phi.tr = c(rep(0,374),1))
est <- est.arma.wge(train, 3, 1, factor = F)

# assess residuals
tswgewrapped::evaluate_residuals(est$res)

# forecast on the test set
f.univat <- fore.aruma.wge(train,est$phi, est$theta, s = 375, n.ahead = length(test), plot = F)

# calculate MSE
mean( (test - f.univat$f)^2 )

# plot the forecast with the test set
plot(seq_along(test) + length(train), test, type = 'l', 
     main = 'Seasonal + ARMA Forecast', xlab = 'Time', ylab = 'Humidity')
lines(seq_along(test) + length(train), f.univat$f, col = 'blue')
```






























































