---
title: "Using tswge"
author: "Stuart Miller"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# tswge

[tswge](https://cran.r-project.org/web/packages/tswge/index.html) is a R package for time series analysis.

```{r}
# load library
library(tswge)

# load some included datasets for work
data(wtcrude) # west texas crude oil prices
data(patemp) # Pennsylvania monthly temperature data
```

## Plotting

### Creating Diagnostic Plots

#### `plotts.sample.wge`

For a quick set of diagonstic plot use `plotts.sample.wge`.
This will plot the realization, the sample autocorrelations, periodogram, and Parzen window.

```{r, diag-plots}
plotts.sample.wge(patemp)
```


## Generating Data

### Simulate Sinusoids with Noise

The function *`gen.sigplusnoise.wge`* generates simulated data based on sinusoids with noise.
It can be used to generate a one- or two-component sinusoid.

Generate a realization from the model `x(t)=coef[1]*cos(2*pi*freq[1]*t+psi[1])+coef[2]*cos(2*pi*freq[2]*t+psi[2])+a(t)`

**Examples**

1. **One component sinusoid**

The following code can be used to simulate 100 data points from the following equation:

`x(t)=3*cos(2*pi*0.05*t+2)+a(t)`

```{r, one-component-sine}
X_t = gen.sigplusnoise.wge(n = 100, coef = c(3,0), freq = c(0.05, 0), psi = c(2, 0))
```

2. **Two component sinusoid**

The following code can be used to simulate 100 data points from the following equation:

`x(t)=3*cos(2*pi*0.05*t+0)+1.5*cos(2*pi*0.35*t+2)+a(t)`

```{r, two-component-sine}
X_t = gen.sigplusnoise.wge(n = 100, coef = c(3, 1.5), freq = c(0.05, 0.35), psi = c(0, 2))
```

